---
title: "Practical Machine Learning"
author: "Cagan KAYA"
date: "Wednesday, October 22, 2014"
output: html_document
---
Executive Summary
----

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.

By this practical machine learning project we discover the relations between barbell lifts with 5 different way and their goal performance

Loading and Cleaning Data
----
```{r,echo=TRUE}
# Loading Files
data.test<-read.csv("pml-testing.csv",na.strings=c("NA",""))
data.training<-read.csv("pml-training.csv",na.strings=c("NA",""))
# Cleaning NA Values From Training and Test Data
tot_na_tr<-sapply(data.training, function(x) sum(is.na(x)))
clean.training.data<-data.training[tot_na_tr==0][,7:60]
tot_na_te<-sapply(data.test, function(x) sum(is.na(x)))
clean.test.data<-data.test[tot_na_te==0][,c(7:60)]
```

Explatory Analysis
---
we explorethe levels of classe variable by plotting
```{r,echo=TRUE}
gr_classe<-data.frame(table(clean.training.data$classe))
plot<-barplot(gr_classe$Freq,col=rainbow(5),names.arg=gr_classe$Var1,ylab="Frequency",xlab="Levels")
```

Modelling
---

#### Creating Partition
We create a partition to data %80 training and %20 validating parts
```{r,echo=TRUE}
library(caret)
library(randomForest)
inputTrain <- createDataPartition(y=clean.training.data$classe, p=0.8, list=FALSE)
tr1 <- clean.training.data[inputTrain,]
tr2 <- clean.training.data[-inputTrain,]
```

#### Building Model
We fit a random forest model
```{r,echo=TRUE,cache=TRUE,}
modelRf <- train(classe ~ ., method = "rf", data = tr1, importance = T, trControl = trainControl(method = "cv", number = 4))
print (modelRf)
```
### Accuracy
we check the model performance on the validation set with in and out of sample tests
In Sample Test
```{r,echo=TRUE,cache=TRUE}
confusionMatrix(predict(modelRf, clean.training.data), clean.training.data$classe)
```

Out Of Sample Test
```{r,echo=TRUE,cache=TRUE}
confusionMatrix(predict(modelRf, tr2), tr2$classe)
```
Results
---
By out of sample test we get %99 accuracy. 
Submitting Answers
---
```{r,echo=TRUE,cache=TRUE}
answers <- predict(modelRf, clean.test.data)
answers
pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}

pml_write_files(answers)
```
